---
title: "Data Cleaning"
author: "Aashka Navale"
date: "2025-04-30"
output: pdf_document
---
# Data Cleaning

### Loading library & datasets

This file has code for the cleaned data. I first load in tidyverse to perform my cleaning operations.

```{r}
library(tidyverse)
```

Here, I'm loading in all four of my csv files. 

```{r}
fertilityRate <- read_csv("adolescentfertilityrate.csv")
totalAdolescents <- read_csv("totaladolescentsoutofschool.csv", locale = locale(encoding = "Latin1"))
femaleAdolescents <- read_csv("femaleadolescentsoutofschool.csv", locale = locale(encoding = "Latin1"))
maleAdolescents <- read_csv("maleadolescentsoutofschool.csv", locale = locale(encoding = "Latin1"))
```

### Cleaning & mutating

In my male adolescents dataset, the country column did not have a name so here I assigned it the name so it would have an identifier.

```{r}
names(maleAdolescents)[1] <- "Country"
```

In my fertility rate dataset, there was an extra row with text that isn't relevant so I removed that row.

```{r}
fertilityRate <- fertilityRate[-c(218), ]
```

For all three of my adolescents datasets, I removed colums 2-9 which were years 1960-1995 since there was an unbearable amount of N/A values. I decided that it wasn't worth it since there simply just wasn't enough to work with. I didn't do this for the fertility rate dataset because it had full data with meaningful values so I could use that well. 

```{r}
femaleAdolescents <- femaleAdolescents[,-c(2:9)]
maleAdolescents <- maleAdolescents[, -c(2:9)]
totalAdolescents <- totalAdolescents[, -c(2:9)]
```

For the following four chunks of code, I mutated to create two new categorical variables in all 4 four datasets. I wanted to create income level and region that categorizes each country into its respective category. 

There are seven regions: North America, Latin America and Caribbean, Middle East and North Africa, Sub-Saharan Africa, Europe and Central Asia, South Asia, and East Asia and Pacific. 

There are four income levels: High income, upper middle income, lower middle income, and low income.

```{r}
fertilityRate2 <- fertilityRate |>
  mutate(Region = case_when(
    
    Country %in% c("United States", "Canada", "Bermuda") ~ "North America",
    
    Country %in% c("Antigua and Barbuda", "Argentina", "Aruba", "Bahamas, The", "Barbados", "Belize", "Bolivia", "Brazil", "British Virgin Islands", "Cayman Islands", "Chile", "Colombia", "Costa Rica", "Cuba", "Curacao", "Dominica", "Dominican Republic", "Ecuador", "El Salvador", "Grenada", "Guatemala", "Guyana", "Haiti", "Honduras", "Jamaica", "Mexico", "Nicaragua", "Panama", "Paraguay", "Peru", "Puerto Rico", "Sint Maarten (Dutch part)", "St. Kitts and Nevis", "St. Lucia", "St. Martin (French part)", "St. Vincent and the Grenadines", "Suriname", "Trinidad and Tobago", "Turks and Caicos Islands", "Uruguay", "Venezuela, RB", "Virgin Islands (U.S.)") ~ "Latin America and Caribbean",
    
    Country %in% c("Algeria", "Bahrain", "Djibouti", "Egypt, Arab Rep.", "Iran, Islamic Rep.", "Iraq", "Israel", "Jordan", "Kuwait", "Lebanon", "Libya", "Morocco", "Oman", "Qatar", "Saudi Arabia", "Syrian Arab Republic", "Tunisia", "United Arab Emirates", "West Bank and Gaza", "Yemen, Rep.") ~ "Middle East and North Africa",
    
    Country %in% c("Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cabo Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Congo, Dem. Rep.", "Congo, Rep.", "Cote d'Ivoire", "Equatorial Guinea", "Eritrea", "Eswatini", "Ethiopia", "Gabon", "Gambia, The", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles", "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania", "Togo", "Uganda", "Zambia", "Zimbabwe") ~ "Sub-Saharan Africa",
    
    Country %in% c("Albania", "Andorra", "Armenia", "Austria", "Azerbaijan", "Belarus", "Belgium", "Bosnia and Herzegovina", "Bulgaria", "Channel Islands", "Croatia", "Cyprus", "Czechia", "Denmark", "Estonia", "Faroe Islands", "Finland", "France", "Georgia", "Germany", "Gibraltar", "Greece", "Greenland", "Hungary", "Iceland", "Ireland", "Isle of Man", "Italy", "Kazakhstan", "Kosovo", "Kyrgyz Republic", "Latvia", "Liechtenstein", "Lithuania", "Luxembourg", "Malta", "Moldova", "Monaco", "Montenegro", "Netherlands", "North Macedonia", "Norway", "Poland", "Portugal", "Romania","Russian Federation", "San Marino", "Serbia", "Slovak Republic", "Slovenia", "Spain", "Sweden", "Switzerland", "Tajikistan", "Turkiye", "Turkmenistan", "Ukraine", "United Kingdom", "Uzbekistan") ~ "Europe and Central Asia",
    
    Country %in% c("India", "Pakistan", "Afghanistan", "Nepal", "Bangladesh", "Bhutan", "Sri Lanka", "Maldives") ~ "South Asia",
    
    Country %in% c("Australia", "American Samoa", "Brunei Darussalam", "Cambodia", "China", "Fiji", "French Polynesia", "Guam", "Hong Kong SAR, China", "Indonesia", "Japan", "Kiribati", "Korea, Dem. People's Rep.", "Korea, Rep.", "Lao PDR", "Macao SAR, China", "Malaysia", "Marshall Islands", "Micronesia, Fed. Sts.", "Mongolia", "Myanmar", "Nauru", "New Caledonia", "New Zealand", "Northern Mariana Islands", "Palau", "Papua New Guinea", "Philippines", "Samoa", "Singapore", "Solomon Islands", "Thailand", "Timor-Leste", "Tonga", "Tuvalu", "Vanuatu", "Viet Nam") ~ "East Asia and Pacific"
    
    ))

fertilityRate2 <- fertilityRate2 |>
  mutate(Income = case_when(
    
    Country %in% c("American Samoa", "Gibraltar", "Palau", "Andorra", "Greece", "Panama", "Antigua and Barbuda", "Greenland", "Poland", "Aruba", "Guam", "Portugal", "Australia", "Guyana", "Puerto Rico", "Austria", "Hong Kong SAR, China", "Qatar", "Bahamas, The", "Hungary", "Romania", "Bahrain", "Iceland", "Russian Federation", "Barbados", "Ireland", "San Marino", "Belgium", "Isle of Man", "Saudi Arabia", "Bermuda", "Israel", "Seychelles", "British Virgin Islands", "Italy", "Singapore", "Brunei Darussalam", "Japan", "Sint Maarten (Dutch part)", "Bulgaria", "Korea, Rep.", "Slovak Republic", "Canada", "Kuwait", "Slovenia", "Cayman Islands", "Latvia", "Spain", "Channel Islands", "Liechtenstein", "St. Kitts and Nevis", "Chile", "Lithuania", "St. Martin (French part)", "Croatia", "Luxembourg", "Sweden", "Curacao", "Macao SAR, China", "Switzerland", "Cyprus", "Malta", "Taiwan, China", "Czechia", "Monaco", "Trinidad and Tobago", "Denmark", "Nauru", "Turks and Caicos Islands", "Estonia", "Netherlands", "United Arab Emirates", "Faroe Islands", "New Caledonia", "United Kingdom", "Finland", "New Zealand", "United States", "France", "Northern Mariana Islands", "Uruguay", "French Polynesia", "Norway", "Virgin Islands (U.S.)", "Germany", "Oman") ~ "High income",
    
    Country %in% c("Albania", "Equatorial Guinea", "Moldova", "Algeria", "Fiji", "Mongolia", "Argentina", "Gabon", "Montenegro", "Armenia", "Georgia", "Namibia", "Azerbaijan", "Grenada", "North Macedonia", "Belarus", "Guatemala", "Paraguay", "Belize", "Indonesia", "Peru", "Bosnia and Herzegovina", "Iran, Islamic Rep.", "Serbia", "Botswana", "Iraq", "South Africa", "Brazil", "Jamaica", "St. Lucia", "China", "Kazakhstan", "St. Vincent and the Grenadines", "Colombia", "Kosovo", "Suriname", "Costa Rica", "Libya", "Thailand", "Cuba", "Malaysia", "Tonga", "Dominica", "Maldives", "Turkiye", "Dominican Republic", "Marshall Islands", "Turkmenistan", "Ecuador", "Mauritius", "Tuvalu", "El Salvador", "Mexico", "Ukraine", "Venezuela, RB") ~ "Upper middle income",
    
    Country %in% c("Angola", "Bangladesh", "Benin", "Bhutan", "Bolivia", "Cabo Verde", "Cambodia", "Cameroon", "Comoros", "Congo, Rep.", "Cote d'Ivoire", "Djibouti", "Egypt, Arab Rep.", "Eswatini", "Ghana", "Guinea", "Haiti", "Honduras", "India", "Jordan", "Kenya", "Kiribati", "Kyrgyz Republic", "Lao PDR", "Lebanon", "Lesotho", "Mauritania", "Micronesia, Fed. Sts.", "Morocco", "Myanmar", "Nepal", "Nicaragua", "Nigeria", "Pakistan", "Papua New Guinea", "Philippines", "Samoa", "Sao Tome and Principe", "Senegal", "Solomon Islands", "Sri Lanka", "Tajikistan", "Tanzania", "Timor-Leste", "Tunisia", "Uzbekistan", "Vanuatu", "Viet Nam", "West Bank and Gaza", "Zambia", "Zimbabwe") ~ "Lower middle income",
    
    Country %in% c("Afghanistan", "Burkina Faso", "Burundi", "Central African Republic", "Chad", "Congo, Dem. Rep.", "Eritrea", "Ethiopia", "Gambia, The", "Guinea-Bissau", "Korea, Dem. People's Rep.", "Liberia", "Madagascar", "Malawi", "Mali", "Mozambique", "Niger", "Rwanda", "Sierra Leone", "Somalia", "South Sudan", "Sudan", "Syrian Arab Republic", "Togo", "Uganda", "Yemen, Rep.") ~ "Low income"
    
  ))
```

```{r}
femaleAdolescents2 <- femaleAdolescents |>
  mutate(Region = case_when(
    
    Country %in% c("United States", "Canada", "Bermuda") ~ "North America",
    
    Country %in% c("Antigua and Barbuda", "Argentina", "Aruba", "Bahamas, The", "Barbados", "Belize", "Bolivia", "Brazil", "British Virgin Islands", "Cayman Islands", "Chile", "Colombia", "Costa Rica", "Cuba", "Curacao", "Dominica", "Dominican Republic", "Ecuador", "El Salvador", "Grenada", "Guatemala", "Guyana", "Haiti", "Honduras", "Jamaica", "Mexico", "Nicaragua", "Panama", "Paraguay", "Peru", "Puerto Rico", "Sint Maarten (Dutch part)", "St. Kitts and Nevis", "St. Lucia", "St. Martin (French part)", "St. Vincent and the Grenadines", "Suriname", "Trinidad and Tobago", "Turks and Caicos Islands", "Uruguay", "Venezuela, RB", "Virgin Islands (U.S.)") ~ "Latin America and Caribbean",
    
    Country %in% c("Algeria", "Bahrain", "Djibouti", "Egypt, Arab Rep.", "Iran, Islamic Rep.", "Iraq", "Israel", "Jordan", "Kuwait", "Lebanon", "Libya", "Morocco", "Oman", "Qatar", "Saudi Arabia", "Syrian Arab Republic", "Tunisia", "United Arab Emirates", "West Bank and Gaza", "Yemen, Rep.") ~ "Middle East and North Africa",
    
    Country %in% c("Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cabo Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Congo, Dem. Rep.", "Congo, Rep.", "Cote d'Ivoire", "Equatorial Guinea", "Eritrea", "Eswatini", "Ethiopia", "Gabon", "Gambia, The", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles", "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania", "Togo", "Uganda", "Zambia", "Zimbabwe") ~ "Sub-Saharan Africa",
    
    Country %in% c("Albania", "Andorra", "Armenia", "Austria", "Azerbaijan", "Belarus", "Belgium", "Bosnia and Herzegovina", "Bulgaria", "Channel Islands", "Croatia", "Cyprus", "Czechia", "Denmark", "Estonia", "Faroe Islands", "Finland", "France", "Georgia", "Germany", "Gibraltar", "Greece", "Greenland", "Hungary", "Iceland", "Ireland", "Isle of Man", "Italy", "Kazakhstan", "Kosovo", "Kyrgyz Republic", "Latvia", "Liechtenstein", "Lithuania", "Luxembourg", "Malta", "Moldova", "Monaco", "Montenegro", "Netherlands", "North Macedonia", "Norway", "Poland", "Portugal", "Romania","Russian Federation", "San Marino", "Serbia", "Slovak Republic", "Slovenia", "Spain", "Sweden", "Switzerland", "Tajikistan", "Turkiye", "Turkmenistan", "Ukraine", "United Kingdom", "Uzbekistan") ~ "Europe and Central Asia",
    
    Country %in% c("India", "Pakistan", "Afghanistan", "Nepal", "Bangladesh", "Bhutan", "Sri Lanka", "Maldives") ~ "South Asia",
    
    Country %in% c("Australia", "American Samoa", "Brunei Darussalam", "Cambodia", "China", "Fiji", "French Polynesia", "Guam", "Hong Kong SAR, China", "Indonesia", "Japan", "Kiribati", "Korea, Dem. People's Rep.", "Korea, Rep.", "Lao PDR", "Macao SAR, China", "Malaysia", "Marshall Islands", "Micronesia, Fed. Sts.", "Mongolia", "Myanmar", "Nauru", "New Caledonia", "New Zealand", "Northern Mariana Islands", "Palau", "Papua New Guinea", "Philippines", "Samoa", "Singapore", "Solomon Islands", "Thailand", "Timor-Leste", "Tonga", "Tuvalu", "Vanuatu", "Viet Nam") ~ "East Asia and Pacific"
    
    ))

femaleAdolescents2 <- femaleAdolescents2 |>
  mutate(Income = case_when(
    
    Country %in% c("American Samoa", "Gibraltar", "Palau", "Andorra", "Greece", "Panama", "Antigua and Barbuda", "Greenland", "Poland", "Aruba", "Guam", "Portugal", "Australia", "Guyana", "Puerto Rico", "Austria", "Hong Kong SAR, China", "Qatar", "Bahamas, The", "Hungary", "Romania", "Bahrain", "Iceland", "Russian Federation", "Barbados", "Ireland", "San Marino", "Belgium", "Isle of Man", "Saudi Arabia", "Bermuda", "Israel", "Seychelles", "British Virgin Islands", "Italy", "Singapore", "Brunei Darussalam", "Japan", "Sint Maarten (Dutch part)", "Bulgaria", "Korea, Rep.", "Slovak Republic", "Canada", "Kuwait", "Slovenia", "Cayman Islands", "Latvia", "Spain", "Channel Islands", "Liechtenstein", "St. Kitts and Nevis", "Chile", "Lithuania", "St. Martin (French part)", "Croatia", "Luxembourg", "Sweden", "Curacao", "Macao SAR, China", "Switzerland", "Cyprus", "Malta", "Taiwan, China", "Czechia", "Monaco", "Trinidad and Tobago", "Denmark", "Nauru", "Turks and Caicos Islands", "Estonia", "Netherlands", "United Arab Emirates", "Faroe Islands", "New Caledonia", "United Kingdom", "Finland", "New Zealand", "United States", "France", "Northern Mariana Islands", "Uruguay", "French Polynesia", "Norway", "Virgin Islands (U.S.)", "Germany", "Oman") ~ "High income",
    
    Country %in% c("Albania", "Equatorial Guinea", "Moldova", "Algeria", "Fiji", "Mongolia", "Argentina", "Gabon", "Montenegro", "Armenia", "Georgia", "Namibia", "Azerbaijan", "Grenada", "North Macedonia", "Belarus", "Guatemala", "Paraguay", "Belize", "Indonesia", "Peru", "Bosnia and Herzegovina", "Iran, Islamic Rep.", "Serbia", "Botswana", "Iraq", "South Africa", "Brazil", "Jamaica", "St. Lucia", "China", "Kazakhstan", "St. Vincent and the Grenadines", "Colombia", "Kosovo", "Suriname", "Costa Rica", "Libya", "Thailand", "Cuba", "Malaysia", "Tonga", "Dominica", "Maldives", "Turkiye", "Dominican Republic", "Marshall Islands", "Turkmenistan", "Ecuador", "Mauritius", "Tuvalu", "El Salvador", "Mexico", "Ukraine", "Venezuela, RB") ~ "Upper middle income",
    
    Country %in% c("Angola", "Bangladesh", "Benin", "Bhutan", "Bolivia", "Cabo Verde", "Cambodia", "Cameroon", "Comoros", "Congo, Rep.", "Cote d'Ivoire", "Djibouti", "Egypt, Arab Rep.", "Eswatini", "Ghana", "Guinea", "Haiti", "Honduras", "India", "Jordan", "Kenya", "Kiribati", "Kyrgyz Republic", "Lao PDR", "Lebanon", "Lesotho", "Mauritania", "Micronesia, Fed. Sts.", "Morocco", "Myanmar", "Nepal", "Nicaragua", "Nigeria", "Pakistan", "Papua New Guinea", "Philippines", "Samoa", "Sao Tome and Principe", "Senegal", "Solomon Islands", "Sri Lanka", "Tajikistan", "Tanzania", "Timor-Leste", "Tunisia", "Uzbekistan", "Vanuatu", "Viet Nam", "West Bank and Gaza", "Zambia", "Zimbabwe") ~ "Lower middle income",
    
    Country %in% c("Afghanistan", "Burkina Faso", "Burundi", "Central African Republic", "Chad", "Congo, Dem. Rep.", "Eritrea", "Ethiopia", "Gambia, The", "Guinea-Bissau", "Korea, Dem. People's Rep.", "Liberia", "Madagascar", "Malawi", "Mali", "Mozambique", "Niger", "Rwanda", "Sierra Leone", "Somalia", "South Sudan", "Sudan", "Syrian Arab Republic", "Togo", "Uganda", "Yemen, Rep.") ~ "Low income"
    
  ))
```

```{r}
maleAdolescents2 <- maleAdolescents |>
  mutate(Region = case_when(
    
    Country %in% c("United States", "Canada", "Bermuda") ~ "North America",
    
    Country %in% c("Antigua and Barbuda", "Argentina", "Aruba", "Bahamas, The", "Barbados", "Belize", "Bolivia", "Brazil", "British Virgin Islands", "Cayman Islands", "Chile", "Colombia", "Costa Rica", "Cuba", "Curacao", "Dominica", "Dominican Republic", "Ecuador", "El Salvador", "Grenada", "Guatemala", "Guyana", "Haiti", "Honduras", "Jamaica", "Mexico", "Nicaragua", "Panama", "Paraguay", "Peru", "Puerto Rico", "Sint Maarten (Dutch part)", "St. Kitts and Nevis", "St. Lucia", "St. Martin (French part)", "St. Vincent and the Grenadines", "Suriname", "Trinidad and Tobago", "Turks and Caicos Islands", "Uruguay", "Venezuela, RB", "Virgin Islands (U.S.)") ~ "Latin America and Caribbean",
    
    Country %in% c("Algeria", "Bahrain", "Djibouti", "Egypt, Arab Rep.", "Iran, Islamic Rep.", "Iraq", "Israel", "Jordan", "Kuwait", "Lebanon", "Libya", "Morocco", "Oman", "Qatar", "Saudi Arabia", "Syrian Arab Republic", "Tunisia", "United Arab Emirates", "West Bank and Gaza", "Yemen, Rep.") ~ "Middle East and North Africa",
    
    Country %in% c("Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cabo Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Congo, Dem. Rep.", "Congo, Rep.", "Cote d'Ivoire", "Equatorial Guinea", "Eritrea", "Eswatini", "Ethiopia", "Gabon", "Gambia, The", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles", "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania", "Togo", "Uganda", "Zambia", "Zimbabwe") ~ "Sub-Saharan Africa",
    
    Country %in% c("Albania", "Andorra", "Armenia", "Austria", "Azerbaijan", "Belarus", "Belgium", "Bosnia and Herzegovina", "Bulgaria", "Channel Islands", "Croatia", "Cyprus", "Czechia", "Denmark", "Estonia", "Faroe Islands", "Finland", "France", "Georgia", "Germany", "Gibraltar", "Greece", "Greenland", "Hungary", "Iceland", "Ireland", "Isle of Man", "Italy", "Kazakhstan", "Kosovo", "Kyrgyz Republic", "Latvia", "Liechtenstein", "Lithuania", "Luxembourg", "Malta", "Moldova", "Monaco", "Montenegro", "Netherlands", "North Macedonia", "Norway", "Poland", "Portugal", "Romania","Russian Federation", "San Marino", "Serbia", "Slovak Republic", "Slovenia", "Spain", "Sweden", "Switzerland", "Tajikistan", "Turkiye", "Turkmenistan", "Ukraine", "United Kingdom", "Uzbekistan") ~ "Europe and Central Asia",
    
    Country %in% c("India", "Pakistan", "Afghanistan", "Nepal", "Bangladesh", "Bhutan", "Sri Lanka", "Maldives") ~ "South Asia",
    
    Country %in% c("Australia", "American Samoa", "Brunei Darussalam", "Cambodia", "China", "Fiji", "French Polynesia", "Guam", "Hong Kong SAR, China", "Indonesia", "Japan", "Kiribati", "Korea, Dem. People's Rep.", "Korea, Rep.", "Lao PDR", "Macao SAR, China", "Malaysia", "Marshall Islands", "Micronesia, Fed. Sts.", "Mongolia", "Myanmar", "Nauru", "New Caledonia", "New Zealand", "Northern Mariana Islands", "Palau", "Papua New Guinea", "Philippines", "Samoa", "Singapore", "Solomon Islands", "Thailand", "Timor-Leste", "Tonga", "Tuvalu", "Vanuatu", "Viet Nam") ~ "East Asia and Pacific"
    
    ))

maleAdolescents2 <- maleAdolescents2 |>
  mutate(Income = case_when(
    
    Country %in% c("American Samoa", "Gibraltar", "Palau", "Andorra", "Greece", "Panama", "Antigua and Barbuda", "Greenland", "Poland", "Aruba", "Guam", "Portugal", "Australia", "Guyana", "Puerto Rico", "Austria", "Hong Kong SAR, China", "Qatar", "Bahamas, The", "Hungary", "Romania", "Bahrain", "Iceland", "Russian Federation", "Barbados", "Ireland", "San Marino", "Belgium", "Isle of Man", "Saudi Arabia", "Bermuda", "Israel", "Seychelles", "British Virgin Islands", "Italy", "Singapore", "Brunei Darussalam", "Japan", "Sint Maarten (Dutch part)", "Bulgaria", "Korea, Rep.", "Slovak Republic", "Canada", "Kuwait", "Slovenia", "Cayman Islands", "Latvia", "Spain", "Channel Islands", "Liechtenstein", "St. Kitts and Nevis", "Chile", "Lithuania", "St. Martin (French part)", "Croatia", "Luxembourg", "Sweden", "Curacao", "Macao SAR, China", "Switzerland", "Cyprus", "Malta", "Taiwan, China", "Czechia", "Monaco", "Trinidad and Tobago", "Denmark", "Nauru", "Turks and Caicos Islands", "Estonia", "Netherlands", "United Arab Emirates", "Faroe Islands", "New Caledonia", "United Kingdom", "Finland", "New Zealand", "United States", "France", "Northern Mariana Islands", "Uruguay", "French Polynesia", "Norway", "Virgin Islands (U.S.)", "Germany", "Oman") ~ "High income",
    
    Country %in% c("Albania", "Equatorial Guinea", "Moldova", "Algeria", "Fiji", "Mongolia", "Argentina", "Gabon", "Montenegro", "Armenia", "Georgia", "Namibia", "Azerbaijan", "Grenada", "North Macedonia", "Belarus", "Guatemala", "Paraguay", "Belize", "Indonesia", "Peru", "Bosnia and Herzegovina", "Iran, Islamic Rep.", "Serbia", "Botswana", "Iraq", "South Africa", "Brazil", "Jamaica", "St. Lucia", "China", "Kazakhstan", "St. Vincent and the Grenadines", "Colombia", "Kosovo", "Suriname", "Costa Rica", "Libya", "Thailand", "Cuba", "Malaysia", "Tonga", "Dominica", "Maldives", "Turkiye", "Dominican Republic", "Marshall Islands", "Turkmenistan", "Ecuador", "Mauritius", "Tuvalu", "El Salvador", "Mexico", "Ukraine", "Venezuela, RB") ~ "Upper middle income",
    
    Country %in% c("Angola", "Bangladesh", "Benin", "Bhutan", "Bolivia", "Cabo Verde", "Cambodia", "Cameroon", "Comoros", "Congo, Rep.", "Cote d'Ivoire", "Djibouti", "Egypt, Arab Rep.", "Eswatini", "Ghana", "Guinea", "Haiti", "Honduras", "India", "Jordan", "Kenya", "Kiribati", "Kyrgyz Republic", "Lao PDR", "Lebanon", "Lesotho", "Mauritania", "Micronesia, Fed. Sts.", "Morocco", "Myanmar", "Nepal", "Nicaragua", "Nigeria", "Pakistan", "Papua New Guinea", "Philippines", "Samoa", "Sao Tome and Principe", "Senegal", "Solomon Islands", "Sri Lanka", "Tajikistan", "Tanzania", "Timor-Leste", "Tunisia", "Uzbekistan", "Vanuatu", "Viet Nam", "West Bank and Gaza", "Zambia", "Zimbabwe") ~ "Lower middle income",
    
    Country %in% c("Afghanistan", "Burkina Faso", "Burundi", "Central African Republic", "Chad", "Congo, Dem. Rep.", "Eritrea", "Ethiopia", "Gambia, The", "Guinea-Bissau", "Korea, Dem. People's Rep.", "Liberia", "Madagascar", "Malawi", "Mali", "Mozambique", "Niger", "Rwanda", "Sierra Leone", "Somalia", "South Sudan", "Sudan", "Syrian Arab Republic", "Togo", "Uganda", "Yemen, Rep.") ~ "Low income"
    
  ))
```

```{r}
totalAdolescents2 <- totalAdolescents |>
  mutate(Region = case_when(
    
    Country %in% c("United States", "Canada", "Bermuda") ~ "North America",
    
    Country %in% c("Antigua and Barbuda", "Argentina", "Aruba", "Bahamas, The", "Barbados", "Belize", "Bolivia", "Brazil", "British Virgin Islands", "Cayman Islands", "Chile", "Colombia", "Costa Rica", "Cuba", "Curacao", "Dominica", "Dominican Republic", "Ecuador", "El Salvador", "Grenada", "Guatemala", "Guyana", "Haiti", "Honduras", "Jamaica", "Mexico", "Nicaragua", "Panama", "Paraguay", "Peru", "Puerto Rico", "Sint Maarten (Dutch part)", "St. Kitts and Nevis", "St. Lucia", "St. Martin (French part)", "St. Vincent and the Grenadines", "Suriname", "Trinidad and Tobago", "Turks and Caicos Islands", "Uruguay", "Venezuela, RB", "Virgin Islands (U.S.)") ~ "Latin America and Caribbean",
    
    Country %in% c("Algeria", "Bahrain", "Djibouti", "Egypt, Arab Rep.", "Iran, Islamic Rep.", "Iraq", "Israel", "Jordan", "Kuwait", "Lebanon", "Libya", "Morocco", "Oman", "Qatar", "Saudi Arabia", "Syrian Arab Republic", "Tunisia", "United Arab Emirates", "West Bank and Gaza", "Yemen, Rep.") ~ "Middle East and North Africa",
    
    Country %in% c("Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cabo Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Congo, Dem. Rep.", "Congo, Rep.", "Cote d'Ivoire", "Equatorial Guinea", "Eritrea", "Eswatini", "Ethiopia", "Gabon", "Gambia, The", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles", "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania", "Togo", "Uganda", "Zambia", "Zimbabwe") ~ "Sub-Saharan Africa",
    
    Country %in% c("Albania", "Andorra", "Armenia", "Austria", "Azerbaijan", "Belarus", "Belgium", "Bosnia and Herzegovina", "Bulgaria", "Channel Islands", "Croatia", "Cyprus", "Czechia", "Denmark", "Estonia", "Faroe Islands", "Finland", "France", "Georgia", "Germany", "Gibraltar", "Greece", "Greenland", "Hungary", "Iceland", "Ireland", "Isle of Man", "Italy", "Kazakhstan", "Kosovo", "Kyrgyz Republic", "Latvia", "Liechtenstein", "Lithuania", "Luxembourg", "Malta", "Moldova", "Monaco", "Montenegro", "Netherlands", "North Macedonia", "Norway", "Poland", "Portugal", "Romania","Russian Federation", "San Marino", "Serbia", "Slovak Republic", "Slovenia", "Spain", "Sweden", "Switzerland", "Tajikistan", "Turkiye", "Turkmenistan", "Ukraine", "United Kingdom", "Uzbekistan") ~ "Europe and Central Asia",
    
    Country %in% c("India", "Pakistan", "Afghanistan", "Nepal", "Bangladesh", "Bhutan", "Sri Lanka", "Maldives") ~ "South Asia",
    
    Country %in% c("Australia", "American Samoa", "Brunei Darussalam", "Cambodia", "China", "Fiji", "French Polynesia", "Guam", "Hong Kong SAR, China", "Indonesia", "Japan", "Kiribati", "Korea, Dem. People's Rep.", "Korea, Rep.", "Lao PDR", "Macao SAR, China", "Malaysia", "Marshall Islands", "Micronesia, Fed. Sts.", "Mongolia", "Myanmar", "Nauru", "New Caledonia", "New Zealand", "Northern Mariana Islands", "Palau", "Papua New Guinea", "Philippines", "Samoa", "Singapore", "Solomon Islands", "Thailand", "Timor-Leste", "Tonga", "Tuvalu", "Vanuatu", "Viet Nam") ~ "East Asia and Pacific"
    
    ))

totalAdolescents2 <- totalAdolescents2 |>
  mutate(Income = case_when(
    
    Country %in% c("American Samoa", "Gibraltar", "Palau", "Andorra", "Greece", "Panama", "Antigua and Barbuda", "Greenland", "Poland", "Aruba", "Guam", "Portugal", "Australia", "Guyana", "Puerto Rico", "Austria", "Hong Kong SAR, China", "Qatar", "Bahamas, The", "Hungary", "Romania", "Bahrain", "Iceland", "Russian Federation", "Barbados", "Ireland", "San Marino", "Belgium", "Isle of Man", "Saudi Arabia", "Bermuda", "Israel", "Seychelles", "British Virgin Islands", "Italy", "Singapore", "Brunei Darussalam", "Japan", "Sint Maarten (Dutch part)", "Bulgaria", "Korea, Rep.", "Slovak Republic", "Canada", "Kuwait", "Slovenia", "Cayman Islands", "Latvia", "Spain", "Channel Islands", "Liechtenstein", "St. Kitts and Nevis", "Chile", "Lithuania", "St. Martin (French part)", "Croatia", "Luxembourg", "Sweden", "Curacao", "Macao SAR, China", "Switzerland", "Cyprus", "Malta", "Taiwan, China", "Czechia", "Monaco", "Trinidad and Tobago", "Denmark", "Nauru", "Turks and Caicos Islands", "Estonia", "Netherlands", "United Arab Emirates", "Faroe Islands", "New Caledonia", "United Kingdom", "Finland", "New Zealand", "United States", "France", "Northern Mariana Islands", "Uruguay", "French Polynesia", "Norway", "Virgin Islands (U.S.)", "Germany", "Oman") ~ "High income",
    
    Country %in% c("Albania", "Equatorial Guinea", "Moldova", "Algeria", "Fiji", "Mongolia", "Argentina", "Gabon", "Montenegro", "Armenia", "Georgia", "Namibia", "Azerbaijan", "Grenada", "North Macedonia", "Belarus", "Guatemala", "Paraguay", "Belize", "Indonesia", "Peru", "Bosnia and Herzegovina", "Iran, Islamic Rep.", "Serbia", "Botswana", "Iraq", "South Africa", "Brazil", "Jamaica", "St. Lucia", "China", "Kazakhstan", "St. Vincent and the Grenadines", "Colombia", "Kosovo", "Suriname", "Costa Rica", "Libya", "Thailand", "Cuba", "Malaysia", "Tonga", "Dominica", "Maldives", "Turkiye", "Dominican Republic", "Marshall Islands", "Turkmenistan", "Ecuador", "Mauritius", "Tuvalu", "El Salvador", "Mexico", "Ukraine", "Venezuela, RB") ~ "Upper middle income",
    
    Country %in% c("Angola", "Bangladesh", "Benin", "Bhutan", "Bolivia", "Cabo Verde", "Cambodia", "Cameroon", "Comoros", "Congo, Rep.", "Cote d'Ivoire", "Djibouti", "Egypt, Arab Rep.", "Eswatini", "Ghana", "Guinea", "Haiti", "Honduras", "India", "Jordan", "Kenya", "Kiribati", "Kyrgyz Republic", "Lao PDR", "Lebanon", "Lesotho", "Mauritania", "Micronesia, Fed. Sts.", "Morocco", "Myanmar", "Nepal", "Nicaragua", "Nigeria", "Pakistan", "Papua New Guinea", "Philippines", "Samoa", "Sao Tome and Principe", "Senegal", "Solomon Islands", "Sri Lanka", "Tajikistan", "Tanzania", "Timor-Leste", "Tunisia", "Uzbekistan", "Vanuatu", "Viet Nam", "West Bank and Gaza", "Zambia", "Zimbabwe") ~ "Lower middle income",
    
    Country %in% c("Afghanistan", "Burkina Faso", "Burundi", "Central African Republic", "Chad", "Congo, Dem. Rep.", "Eritrea", "Ethiopia", "Gambia, The", "Guinea-Bissau", "Korea, Dem. People's Rep.", "Liberia", "Madagascar", "Malawi", "Mali", "Mozambique", "Niger", "Rwanda", "Sierra Leone", "Somalia", "South Sudan", "Sudan", "Syrian Arab Republic", "Togo", "Uganda", "Yemen, Rep.") ~ "Low income"
    
  ))
```

### Exporting files

These final chunks of code is to export the cleaned data into my computer. 

```{r}
df1 <- data.frame(fertilityRate2)

write_csv(df1, "/Users/aashkanavale/Desktop/MC Spring 25/DATA205/Project/Datasets/Cleaned Data/fertilityRate.csv")
```

```{r}
df2 <- data.frame(femaleAdolescents2)

write_csv(df2, "/Users/aashkanavale/Desktop/MC Spring 25/DATA205/Project/Datasets/Cleaned Data/femaleAdolescents.csv")
```

```{r}
df3 <- data.frame(maleAdolescents2)

write_csv(df3, "/Users/aashkanavale/Desktop/MC Spring 25/DATA205/Project/Datasets/Cleaned Data/maleAdolescents.csv")
```

```{r}
df4 <- data.frame(totalAdolescents2)

write_csv(df4, "/Users/aashkanavale/Desktop/MC Spring 25/DATA205/Project/Datasets/Cleaned Data/totalAdolescents.csv")
```

