---
title: "Final Analyses"
author: "Aashka Navale"
date: "2025-05-06"
output: html_document
---

# Analysis Code

### Loading libraries and datasets

This file contains my final data products for my key research questions. Once again, I'll load in my necessary libraries. I will be using highcharter to creative interactive plots. 

```{r}
library(tidyverse)
library(highcharter)
```

```{r}
fertilityRate <- read_csv("/Users/aashkanavale/Desktop/MC Spring 25/DATA205/Project/Datasets/Cleaned Data/fertilityRate.csv")
femaleAdolescents <- read_csv("/Users/aashkanavale/Desktop/MC Spring 25/DATA205/Project/Datasets/Cleaned Data/femaleAdolescents.csv")
maleAdolescents <- read_csv("/Users/aashkanavale/Desktop/MC Spring 25/DATA205/Project/Datasets/Cleaned Data/maleAdolescents.csv")
totalAdolescents <- read_csv("/Users/aashkanavale/Desktop/MC Spring 25/DATA205/Project/Datasets/Cleaned Data/totalAdolescents.csv")
```

Here I'm renaming my variables because when R saved them to my computer, there was an 'X' in front of the years so I'm simply renaming them to get rid of that. 

```{r}
names(fertilityRate) <- c("Country", "1960", "1965", "1970", "1975", "1980", "1985", "1990", "1995", "2000", "2005", "2010", "2015", "2020", "Region", "Income")
names(femaleAdolescents) <- c("Country", "2000", "2005", "2010", "2015", "2020", "Region", "Income")
names(maleAdolescents) <- c("Country", "2000", "2005", "2010", "2015", "2020", "Region", "Income")
names(totalAdolescents) <- c("Country", "2000", "2005", "2010", "2015", "2020", "Region", "Income")
```

## 1.0. Is there a correlation between adolescent fertility rates and the percentage of female adolescents out of school across different world regions?

For my first main question, I want to find a relationship between adolescent fertility rates and the percentage of female adolescents out of school. I also wanted to see if a particular region tends to have a higher/lower fertility rate than others.

```{r}
# reformatting to long format using numeric year columns
female_long <- femaleAdolescents |>
  pivot_longer(cols = matches("^\\d{4}$"), names_to = "Year", values_to = "FemaleOutOfSchool") |>
  mutate(Year = as.integer(Year))

fertility_long <- fertilityRate |>
  pivot_longer(cols = matches("^\\d{4}$"), names_to = "Year", values_to = "AdolescentFertilityRate") |>
  mutate(Year = as.integer(Year))

# merging datasets
merged_df <- female_long |>
  inner_join(fertility_long, by = c("Country", "Year", "Region", "Income"))

# cleaning n/a values
merged_df <- merged_df |>
  filter(FemaleOutOfSchool != "..") |>
  mutate(FemaleOutOfSchool = as.numeric(FemaleOutOfSchool),
         AdolescentFertilityRate = as.numeric(AdolescentFertilityRate)) |>
  drop_na()


palette <- c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51")

# series per year
years <- sort(unique(merged_df$Year))

series_list <- merged_df |>
  group_by(Year) |>
  group_split() |>
  lapply(function(df_year) {
    year <- df_year$Year[1]
    year_index <- which(years == year)
    color <- palette[(year_index - 1) %% length(palette) + 1] 

    list(name = as.character(year),
         data = list_parse(df_year |>
                             mutate(name = Country) |>
                             select(x = AdolescentFertilityRate, y = FemaleOutOfSchool, name)),
         type = "scatter",
         color = color, 
         marker = list(radius = 5),
         tooltip = list(pointFormat = "<b>{point.name}</b><br/>Fertility: {point.x}<br/>Out of School: {point.y}%"))})
```

```{r}
# scatterplot using highcharter
highchart() |>
  hc_chart(type = "scatter") |>
  hc_title(text = "Adolescent Fertility vs. Female Adolescents Out of School") |>
  hc_xAxis(title = list(text = "Adolescent Fertility Rate (births/1000 girls)")) |>
  hc_yAxis(title = list(text = "% of Female Adolescents Out of School")) |>
  hc_add_series_list(series_list) |>
  hc_legend(enabled = TRUE)
```

This scatterplot shows the relationship between adolescent fertility rates and the percentage of female adolescents out of school across different countries from 2000 to 2020. Each point stands for a country in a specific year, and the colors and shapes differentiate the years. Overall, there is a clear trend. Countries with higher adolescent fertility rates tend to have more girls out of school. Most data points are clustered in the bottom left of the graph, suggesting that many countries have relatively low fertility and better school attendance. Over time, especially by 2020, more countries appear to have lower fertility rates, showing progress in reproductive health.

### 1.1. How does this relationship differ between Sub-Saharan Africa and the global average since 2000?

For this next sub-question, I wanted to compare Sub-Saharan Africa to the rest of the world. I averaged out the region and global data.

```{r}
summary_df <- merged_df |> 
  filter(Year >= 2000) |>
  group_by(Region = ifelse(Region == "Sub-Saharan Africa", "Sub-Saharan Africa", "Global"), Year) |>
  summarise(Fertility = mean(AdolescentFertilityRate, na.rm = TRUE),
            Dropout = mean(FemaleOutOfSchool, na.rm = TRUE),
            .groups = "drop") |>
  group_by(Region, Year) |>
  summarise(Fertility = mean(Fertility),
            Dropout = mean(Dropout),
            .groups = "drop")

# splitting out Sub-Saharan Africa and global
ssa <- summary_df |> filter(Region == "Sub-Saharan Africa")
global <- merged_df |>
  group_by(Year) |>
  summarise(Fertility = mean(AdolescentFertilityRate, na.rm = TRUE),
            Dropout = mean(FemaleOutOfSchool, na.rm = TRUE),
            .groups = "drop")

# dual-axis plot using highcharter
highchart() |>
  hc_title(text = "Adolescent Fertility vs. Female Dropout Rate (Sub-Saharan Africa vs Global)") |>
  hc_xAxis(categories = unique(global$Year)) |>
  hc_yAxis_multiples(list(title = list(text = "Fertility Rate"), opposite = FALSE),
                     list(title = list(text = "% Out of School"), opposite = TRUE)) |>
  hc_add_series(name = "SSA Fertility", data = ssa$Fertility, type = "line", yAxis = 0, color = "#006d77") |>
  hc_add_series(name = "SSA Dropout", data = ssa$Dropout, type = "line", yAxis = 1, color = "#83c5be") |>
  hc_add_series(name = "Global Fertility", data = global$Fertility, type = "line", yAxis = 0, dashStyle = "ShortDot", color = "#33658a") |>
  hc_add_series(name = "Global Dropout", data = global$Dropout, type = "line", yAxis = 1, dashStyle = "ShortDot", color = "#86bbd8") |>
  hc_tooltip(shared = TRUE,
             valueDecimals = 2) |>
  hc_tooltip(shared = TRUE) |>
  hc_legend(enabled = TRUE)
```

These results are astounding. Sub-Saharan Africa has a significantly larger average fertility and dropout rate than the rest of the world, nearly double. It looks like there's a decline occurring in recent years and based on educated inferences, it will continue to go down as time passes. The highest peak seems to be in 2005. I would have liked to explore data from 1960 as well but it was so sparse that I stated it unworkable. 

### 1.2. Are changes in adolescent fertility associated with increases or decreases in school dropout rates among adolescent girls?

```{r}
# reformatting to long format
fertility_long <- fertilityRate |>
  pivot_longer(cols = matches("^[0-9]{4}$"),
               names_to = "Year",
               values_to = "FertilityRate") |>
  mutate(Year = as.integer(Year),
         FertilityRate = as.numeric(FertilityRate))

female_long <- femaleAdolescents |>
  pivot_longer(cols = matches("^[0-9]{4}$"),
               names_to = "Year",
               values_to = "FemaleDropoutRate") |>
  mutate(Year = as.integer(Year),
         FemaleDropoutRate = as.numeric(FemaleDropoutRate))

# merging datasets
merged <- inner_join(fertility_long, female_long,
                     by = c("Country", "Year", "Region", "Income")) |>
  drop_na(FertilityRate, FemaleDropoutRate)

# plotting using ggplot2
ggplot(merged, aes(x = FertilityRate, y = FemaleDropoutRate, color = Region)) +
  geom_point(alpha = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  scale_color_manual(values = c("#264653", "#287271", "#2a9d8f", "#e9c46a", "#f4a261", "#ee8959", "#e76f51"))+
  labs(title = "Adolescent Fertility vs. Female School Dropout Rates",
       x = "Adolescent Fertility Rate (births per 1,000 girls)",
       y = "Female School Dropout Rate (%)") +
  theme_minimal() +
  theme(legend.position = "right")
```

This is similar to graph 1.0, but the only difference is that it's colored by region instead of year. Now it's much easier to see the spread of the regions. Once again, Sub-Saharan African countries prevail in the most amount of countries with the higheset fertility rate and dropout rate. Overall, there is a clear trend, as shown by the regression line. Countries with higher adolescent fertility rates tend to have more girls out of school. Most data points are clustered in the bottom left of the graph, suggesting that many countries have relatively low fertility and better school attendance.

### 2.0. How does national income level affect adolescents’ school attendance rates?

For my second main question, I wanted to incorporate income into fertility rate and adolescents out of school. I chose to facet wrap a boxplot to compare the year 2000 vs 2020. 

```{r}
# filtering for 2000 and 2020
comparison <- female_long |>
  filter(Year %in% c(2000, 2020)) |>
  drop_na(FemaleDropoutRate)

# manually reordering income levels
comparison$Income <- factor(comparison$Income, levels = c("Low income", "Lower middle income", "Upper middle income", "High income"))

# creating faceted boxplot using ggplot
ggplot(comparison, aes(x = Income, y = FemaleDropoutRate, fill = Income)) +
  geom_boxplot(outlier.alpha = 0.3, alpha = 0.8) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 3, color = "black") +
  facet_wrap(~ Year) +
  scale_fill_manual(values = c(
    "Low income" = "#bfd7ea",
    "Lower middle income" = "#0b3954",
    "Upper middle income" = "#a8dadc",
    "High income" = "#e63946")) +
  labs(title = "Female Dropout Rates by Income Level: 2000 vs 2020",
       x = "Income Group",
       y = "Female Dropout Rate (%)") +
  theme_light() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))
```

This boxplot compares female dropout rates across income groups in the years 2000 and 2020. In both years, dropout rates are highest in low-income countries and decrease as income increases. Over time, there is a noticeable decrease in dropout rates across all income groups, with the most significant drop in low-income and lower-middle-income countries. High-income countries consistently show the lowest dropout rates. The black dots represent the average dropout rate for each group, which also shows improvement from 2000 to 2020. 

### 2.1. Within Europe and Central Asia, is there a consistent trend in school attendance across income groups?

For this sub-question, I wanted to look specifically into Europe and Central Asia since they are the highest income region who show the lowest fertility rates.

```{r}
# ordering income levels
ECA$Income <- factor(ECA$Income, levels = c("Lower middle income", "Upper middle income", "High income"))

# boxplot using ggplot with custom fill colors
ggplot(ECA, aes(x = Income, y = FemaleDropoutRate, fill = Income)) +
  geom_boxplot(outlier.alpha = 0.3, alpha = 0.8) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 3, color = "black") +
  scale_fill_manual(values = c(
    "Lower middle income" = "#dd6e42",
    "Upper middle income" = "#e8dab2",
    "High income" = "#4f6d7a")) +
  labs(title = "Female Dropout Rates by Income Level in Europe & Central Asia in 2020",
       x = "Income Group",
       y = "Female Dropout Rate (%)") +
  theme_minimal() +
  theme(legend.position = "none")
```

This boxplot shows female dropout rates in 2020 across different income levels within the Europe and Central Asia region. The chart includes lower-middle, upper-middle, and high-income countries. If you noticed, low-income countries are not shown because none are a part of this region. Overall, high-income countries had the lowest dropout rates with little variation. Upper-middle-income countries showed a wider range of dropout rates and a slightly higher average compared to lower-middle-income countries. This suggests that while income level generally relates to lower dropout rates, other factors play a part in outcomes.

### 2.2. How does female dropout rates compare to male dropout rates?

For this sub-question, I wanted to look for male adolescents this time and see how their dropout rates compare with female adolescents. 

```{r}
# converting to numeric
female_long <- female_long |>
  mutate(FemaleDropoutRate = as.numeric(FemaleDropoutRate))

# reshaping male dataset
male_long <- maleAdolescents |>
  pivot_longer(cols = matches("^[0-9]{4}$"),
               names_to = "Year",
               values_to = "MaleDropoutRate") |>
  mutate(Year = as.integer(Year))

# converting to numeric
male_long <- male_long |>
  mutate(MaleDropoutRate = as.numeric(MaleDropoutRate)) 

# merging by country, year, region, and income
dropout_combined <- inner_join(female_long, male_long, by = c("Country", "Year", "Region", "Income")) |>
  drop_na(FemaleDropoutRate, MaleDropoutRate) |>
  mutate(GenderGap = FemaleDropoutRate - MaleDropoutRate)

# reshaping to long format with gender column
dropout_long <- dropout_combined |>
  select(Country, Region, Income, Year, FemaleDropoutRate, MaleDropoutRate) |>
  pivot_longer(cols = c(FemaleDropoutRate, MaleDropoutRate),
               names_to = "Gender",
               values_to = "DropoutRate") %>%
  mutate(Gender = recode(Gender,
                         "FemaleDropoutRate" = "Female",
                         "MaleDropoutRate" = "Male"))

# manually reordering income levels
dropout_long$Income <- factor(dropout_long$Income, levels = c("Low income", "Lower middle income", "Upper middle income", "High income"))

# bar chart using ggplot
ggplot(dropout_long |> filter(Year == 2020),
       aes(x = Income, y = DropoutRate, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Female vs Male Dropout Rates by Income Level (2020)",
       x = "Income Group", y = "Dropout Rate (%)") +
  theme_minimal()
```

Now, do not immediately trust these results because the data is heavily filled with N/A values. Based off of what we know already, female adolescent dropout rates should be much higher, especially in low income communities, than male dropout rates. In fact, this is the case for most of my visualizations that involve one of the dropout datasets. Since there are so many N/A values, the data was picked at random to whichever country had any data at all. However, a different perspective would tell us that it makes sense that male students would have higher dropout rates for low income groups because families send their sons to work instead of completing an education. 

### 3.0. In 2020, how did the adolescent fertility rate in South Asia compare to East Asia and the Pacific?

I wanted to compare most countries in Asia (since some are a part of the Europe & Central Asia region) by region to see which region has better statistics. 

```{r}
# numeric fertility rates
fertility_long <- fertility_long |>
  mutate(FertilityRate = as.numeric(FertilityRate))

# filtering for 2020 and South Asia & East Asia and the Pacific
fertility2020 <- fertility_long |>
  filter(Year == 2020,
         Region %in% c("South Asia", "East Asia and Pacific")) |>
  drop_na(FertilityRate)

# summarize by region
summary <- fertility2020 |>
  group_by(Region)|>
  summarise(MeanFertility = mean(FertilityRate, na.rm = TRUE),
            MedianFertility = median(FertilityRate, na.rm = TRUE),
            MinFertility = min(FertilityRate),
            MaxFertility = max(FertilityRate),
            .groups = "drop")
summary
```

This table compares adolescent fertility rates between East Asia and Pacific and South Asia. South Asia has a higher average (mean) fertility rate of 38.5 births per 1,000 girls compared to 29.96 in East Asia and Pacific. The median fertility rate is slightly higher in South Asia (30.05) than in East Asia and Pacific (28.8). While both regions have similar maximum fertility values, 79.2 in South Asia and 83.5 in East Asia and Pacific, the minimum values differ more, with East Asia and Pacific having a very low rate of 0.6, compared to 6.4 in South Asia. This shows that there is a greater variation in fertility rates within the East Asia and Pacific region.

### 3.1 How do adolescent fertility rates vary across world regions, and how have they changed over time?

This visualization will be made in Tableau as a dashboard so I'm reshaping data to export from RStudio onto my computer.

```{r}
fertility_long2 <- fertilityRate |>
  pivot_longer(cols = matches("^[0-9]{4}$"),
               names_to = "Year",
               values_to = "FertilityRate") |>
  mutate(Year = as.integer(Year),
         FertilityRate = as.numeric(FertilityRate)) |>
  drop_na(FertilityRate)

fertility_clean <- fertility_long2 |>
  filter(Year >= 1960 & Year <= 2020) |>
  drop_na(Country, Region, FertilityRate)

# exporting data for tableau vizualization
write_csv(fertility_clean, "fertilityTableau.csv")
```
